The Precambrian paleogeography of Laurentia

### 2021-12-11 Manuscript/Laurentia_Paleogeo_Manuscript.tex
update .tex/PDF to indicate author's version rather than preprint

\begin{abstract}
Laurentia is the craton that forms the Precambrian core of North America and was a major continent throughout the majority of the Proterozoic following its amalgamation 1.8 billion years ago. The paleogeographic position of Laurentia is key to the development of reconstructions of Proterozoic paleogeography including the Paleoproterozoic to Mesoproterozoic supercontinent Nuna and latest Mesoproterozoic to Neoproterozoic supercontinent Rodinia. There is a rich record of Precambrian paleomagnetic poles from Laurentia, as well as an extensive and well-documented geologic history of tectonism. These geologic and paleomagnetic records are increasingly better constrained geochronologically and are both key to evaluating and developing paleogeographic models. These data from Laurentia provide strong-support for mobile lid plate tectonic processes operating continuously over the past 2.2 billion years. 
\end{abstract}
%\keywords{First keyword \and Second keyword \and More} % (optional)
\textit{This manuscript is a preprint of the chapter: \vspace{0.1 cm} \\
Swanson-Hysell, N. L. (2021) The Precambrian paleogeography of Laurentia. In: Pesonen, L.J., Salminen, J., Evans, D.A.D., Elming, S.-Å., Veikkolainen, T. (eds.) Ancient Supercontinents and the Paleogeography of the Earth, doi:10.1016/B978-0-12-818533-9.00009-6.}
\vspace{0.3 cm}
  \end{@twocolumnfalse} % Method A for two-column formatting
] % Method A for two-column formatting

@@

%%%%%%%%%%%%% Acknowledgements %%%%%%%%%%%%%
\footnotesize
\subsection*{Acknowledgements}
This work was supported by NSF CAREER Grant EAR-1847277. The manuscript benefited from reviews from Athena Eyster, David Evans, and Lauri Pesonen as well as discussions and manuscript feedback from Francis Macdonald, Yuem Park, Toby Rivers, Sarah Slotznick, Justin Strauss, and Yiming Zhang. Bruce Buffett provided insights on the implications of a stagnant lid regime for the generation of the geomagnetic field. Many participants in the Nordic Paleomagnetism Workshop have contributed to the compilation and evaluation of the pole list utilized herein. Particular acknowledgement goes to David Evans for maintaining and distributing the compiled pole lists. GPlates, and in particular the pyGPlates API, was utilized in this work \citep{Muller2018b}. Figures were made using Matplotlib \citep{Hunter2007a} in conjunction with cartopy \citep{Met-Office2010a} and pmagpy \citep{Tauxe2016a} within an interactive Python environment \citep{Perez2007a}. The chapter text as well as code, data, and reconstructions used in this paper are openly available and licensed for any form of reuse with attribution (CC BY 4.0) in this repository: \url{https://github.com/Swanson-Hysell-Group/Laurentia_Paleogeography} which is also archived on Zenodo (https://doi.org/10.5281/zenodo.5129927).

>>>>

\begin{abstract}
Laurentia is the craton that forms the Precambrian core of North America and was a major continent throughout the majority of the Proterozoic following its amalgamation 1.8 billion years ago. The paleogeographic position of Laurentia is key to the development of reconstructions of Proterozoic paleogeography including the Paleoproterozoic to Mesoproterozoic supercontinent Nuna and latest Mesoproterozoic to Neoproterozoic supercontinent Rodinia. There is a rich record of Precambrian paleomagnetic poles from Laurentia, as well as an extensive and well-documented geologic history of tectonism. These geologic and paleomagnetic records are increasingly better constrained geochronologically and are both key to evaluating and developing paleogeographic models. These data from Laurentia provide strong-support for mobile lid plate tectonic processes operating continuously over the past 2.2 billion years. 
\end{abstract}
%\keywords{First keyword \and Second keyword \and More} % (optional)
\textit{This manuscript is the author's version of the following chapter and is licensed for reuse and distribution under a Creative Commons Attribution (CC BY 4.0) license: \vspace{0.1 cm} \\
Swanson-Hysell, N. L. (2021) The Precambrian paleogeography of Laurentia. In: Pesonen, L.J., Salminen, J., Evans, D.A.D., Elming, S.-Å., Veikkolainen, T. (eds.) Ancient Supercontinents and the Paleogeography of the Earth, doi:10.1016/B978-0-12-818533-9.00009-6.}
\vspace{0.3 cm}
  \end{@twocolumnfalse} % Method A for two-column formatting
] % Method A for two-column formatting

@@

%%%%%%%%%%%%% Acknowledgements %%%%%%%%%%%%%
\footnotesize
\subsection*{Acknowledgements}
This work was supported by NSF CAREER Grant EAR-1847277. The manuscript benefited from reviews from Athena Eyster, David Evans, and Lauri Pesonen as well as discussions and manuscript feedback from Francis Macdonald, Yuem Park, Toby Rivers, Sarah Slotznick, Justin Strauss, and Yiming Zhang. Bruce Buffett provided insights on the implications of a stagnant lid regime for the generation of the geomagnetic field. Many participants in the Nordic Paleomagnetism Workshop have contributed to the compilation and evaluation of the pole list utilized herein. Particular acknowledgement goes to David Evans for maintaining and distributing the compiled pole lists. GPlates, and in particular the pyGPlates API, was utilized in this work \citep{Muller2018b}. Figures were made using Matplotlib \citep{Hunter2007a} in conjunction with cartopy \citep{Met-Office2010a} and pmagpy \citep{Tauxe2016a} within an interactive Python environment \citep{Perez2007a}. The chapter text as well as code, data, and reconstructions used in this paper are openly available and licensed for any form of reuse with attribution (CC BY 4.0) in this repository: \url{https://github.com/Swanson-Hysell-Group/Laurentia_Paleogeography}. It is  also archived on Zenodo: \url{https://doi.org/10.5281/zenodo.5129927}.
